---
title: "Bootstrapped Confidence Interval for Proportions"
author: "Larissa Neilson"
date: "3/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE)

library(tidyverse)
library(boot)
```

### Recreate the data from the 2014 UCSB Report

22 out of 36 surveyed UCSB community members (61%) identifying as nonbinary/genderqueer responded that they had personally experienced exclusionary, offensive, hostile or intimidating conduct.

```{r}
# Create vectors representing the data
did_exp <- rep(1, 22)
did_not_exp <- rep(0, 14)

# Combine vectors; results in: n = 36, 22 = "1" aka did personally experience exclusionary, offensive, hostile or intimidating conduct, and 14 = "0" aka did not experience this
study_data <- c(did_exp, did_not_exp)
```

### Bootstrap the proportion of genderqueer/nonbinary students experiencing exclusionary behavior at USCB

```{r, results = "hide"}
# Create the function
prop_fun <- function (x,i) {sum(x[i])/length(x[i])}

# Create 10,000 bootstrapped samples
studyboot_10k <- boot(study_data,
                      statistic = prop_fun,
                      R = 10000)
studyboot_10k
```

```{r}
# Convert to dataframe
studyboot_10k_df <- data.frame(bs_prop = studyboot_10k$t)

# Histogram of 10k bootstrap sample proportions
ggplot(data = studyboot_10k_df, aes(x = bs_prop)) +
  geom_histogram(bins = 23, fill = "darkslategray") +
  labs(x = "Proportion of genderqueer/nonbinary students who experienced exclusionary behavior at UCSB", y = "Count") +
  theme_minimal()
```

```{r, results = "hide"}
# Calculate the 95% confidence interval
boot.ci(studyboot_10k, conf = 0.95)
```

For a 95% confidence interval, we expect that the true proportion of the genderqueer/nonbinary UCSB students experiencing exclusionary behavior is between 0.44 and 0.77 (bootstrap percentile interval).

#### Data

2014 UCSB Campus Climate Project Final Report (prepared by Rankin & Associates Consulting, available at http://campusclimate.ucop.edu/_common/files/pdf-climate/ucsb-full-report.pdf).
 
